Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-03-21 paper/tex/paper.tex
started work on detection rate trends

\todo{Add text about detection rates}

@@

@@

We illustrate the distribution of detectable DCOs lie on the sensitivity curve in Figure~\ref{fig:dcos_on_sc}. The shape of the distributions are bounded by the edge of the Milky Way and the infrequency of nearby DCOs. We indicate this with lines of constant distance that form as a boundary for the majority of DCOs, where the remaining scatter is due to variations in mass since lines are plotted for the average chirp mass of each respective DCO. The top edge is not a hard cutoff but is present because for any given distance there is a lower probability of drawing higher frequencies (since binaries spend more time at lower frequencies).

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{dcos_on_sc.png}
    \caption{\tom{This plot needs to be updated after SNR fixes}. Density distribution of detectable DCOs plotted over the LISA sensitivity curve \citep{Robson+2019}. We plot the total signal of binaries at their dominant frequency $n f_{\rm orb}$, such that $n$ is the harmonic that produces the most relative gravitational wave luminosity ($n = 2$ for circular binaries). If the density of points is below our lowest contour (2\%) then we plot the points as scatter points, where their sizes corresponds to their STROOPWAFEL weights. In order to explain the edges of the distribution, we additionally plot lines of constant distance. The dotted lines show the signal for a circular binary of average chirp mass, whilst the dashed line shows the signal for a binary with $e = 0.9$}
    \label{fig:dcos_on_sc}
\end{figure*}

\subsection{Results for fiducial model}\label{sec:fiducial_distributions}
We find that, for our fiducial model, a four year LISA mission will detect 26 BHBH, 27 BHNS and 12 NSNS binaries. Increasing the LISA mission length to ten years changes the number of detections to 41, 45 and 19 respectively. We can note that each shows an increase of approximately $\sqrt{10 / 4} \approx 1.6$ as expected since $\rho \propto \sqrt{T_{\rm obs}}$.

@@

\newcolumntype{S}{>{\centering}p{0.75cm}}
\newcolumntype{Z}{>{\centering\arraybackslash}p{0.75cm}}
\begin{table*}[htb]
    \centering
    \caption{The number of detectable binaries in a 4 and 10 year LISA mission for the 15 different model variations and each DCO type. Each value shows the median and the 90\% confidence interval.}
    \begin{tabular}{c|lll|lll}
        \hline
        \multirow{2}{*}{Model} & \multicolumn{3}{c|}{LISA (4 year)} & \multicolumn{3}{c}{LISA (10 year)} \\ \cline{2-7}
         & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} \\
        \hline
        A & \confinv{25.9}{11.1}{13.6} & \confinv{26.7}{11.9}{14.8} & \confinv{11.3}{6.4}{8.0} & \confinv{42.0}{17.3}{17.3} & \confinv{44.5}{17.8}{20.7} & \confinv{19.3}{8.0}{9.7}\\
        B & \confinv{27.7}{12.6}{12.6} & \confinv{12.0}{6.0}{7.5} & \confinv{4.7}{2.7}{3.3} & \confinv{42.7}{17.6}{17.6} & \confinv{19.5}{7.5}{10.5} & \confinv{8.0}{3.3}{4.0}\\
        C & \confinv{20.6}{9.4}{11.3} & \confinv{6.2}{2.8}{3.4} & \confinv{5.5}{2.7}{4.1} & \confinv{33.8}{13.1}{13.1} & \confinv{10.1}{3.9}{3.9} & \confinv{9.6}{4.1}{4.1}\\
        D & \confinv{20.0}{8.3}{10.0} & \confinv{3.2}{1.6}{1.6} & \confinv{14.6}{8.3}{10.4} & \confinv{31.7}{11.7}{11.7} & \confinv{5.1}{1.9}{2.2} & \confinv{22.9}{10.4}{12.5}\\
        E & \confinv{24.4}{12.2}{14.7} & \confinv{6.8}{3.8}{3.8} & \confinv{0.2}{0.1}{0.1} & \confinv{41.5}{17.1}{17.1} & \confinv{10.6}{3.8}{5.3} & \confinv{0.4}{0.1}{0.2}\\
        F & \confinv{20.0}{10.0}{10.0} & \confinv{11.5}{6.6}{8.2} & \confinv{8.1}{3.6}{5.4} & \confinv{30.0}{12.0}{14.0} & \confinv{19.7}{9.9}{9.9} & \confinv{13.6}{5.4}{6.3}\\
        G & \confinv{25.7}{9.9}{13.8} & \confinv{23.3}{10.6}{10.6} & \confinv{26.5}{15.1}{18.9} & \confinv{39.5}{13.8}{15.8} & \confinv{36.1}{12.7}{17.0} & \confinv{45.4}{18.9}{22.7}\\
        H & \confinv{63.6}{31.8}{47.7} & \confinv{37.4}{20.8}{20.8} & \confinv{14.7}{8.4}{10.5} & \confinv{103.4}{39.8}{55.7} & \confinv{62.3}{24.9}{29.1} & \confinv{25.2}{12.6}{12.6}\\
        I & \confinv{21.7}{10.9}{13.0} & \confinv{56.8}{31.6}{31.6} & \confinv{11.0}{5.5}{6.9} & \confinv{32.6}{13.0}{15.2} & \confinv{94.7}{37.9}{44.2} & \confinv{17.9}{8.3}{8.3}\\
        J & \confinv{32.6}{16.3}{19.6} & \confinv{20.3}{9.0}{13.5} & \confinv{11.0}{6.3}{7.8} & \confinv{52.2}{19.6}{22.8} & \confinv{33.8}{11.3}{15.8} & \confinv{18.8}{7.8}{9.4}\\
        K & \confinv{21.4}{9.7}{11.7} & \confinv{30.6}{13.6}{17.0} & \confinv{11.5}{6.6}{8.2} & \confinv{35.0}{13.6}{15.6} & \confinv{51.1}{20.4}{23.8} & \confinv{19.8}{8.2}{9.9}\\
        L & \confinv{24.7}{9.9}{14.8} & \confinv{26.7}{14.8}{17.8} & \confinv{11.3}{6.4}{8.0} & \confinv{42.0}{14.8}{17.3} & \confinv{44.5}{17.8}{20.7} & \confinv{19.3}{8.0}{9.7}\\
        M & \confinv{29.7}{13.2}{16.5} & \confinv{70.6}{39.2}{39.2} & \confinv{20.4}{11.7}{11.7} & \confinv{46.1}{19.8}{23.1} & \confinv{109.8}{39.2}{54.9} & \confinv{32.1}{14.6}{17.5}\\
        N & \confinv{31.6}{15.8}{19.7} & \confinv{117.7}{58.8}{73.6} & \confinv{51.8}{29.6}{29.6} & \confinv{51.3}{23.7}{23.7} & \confinv{191.3}{73.6}{103.0} & \confinv{81.4}{37.0}{44.4}\\
        O & \confinv{30.2}{15.1}{18.9} & \confinv{96.7}{48.3}{60.4} & \confinv{11.4}{6.5}{8.1} & \confinv{49.1}{22.7}{22.7} & \confinv{157.1}{60.4}{84.6} & \confinv{19.5}{9.7}{9.7}\\
        \hline
    \end{tabular}
    \label{tab:detection_rates}
\end{table*}

>>>>

In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections for each model variation. We find that on average, for our fiducial model, a four year LISA mission will detect 26 BHBH, 27 BHNS and 11 NSNS binaries. Increasing the LISA mission length to ten years changes the number of detections to 42, 45 and 19 respectively. The BHBH detection rate is markedly robust across physics variations, with the expected detections in each model staying within 25\% of the fiducial rate (with the exception of model \modOpt{}). Thus even if there are changes in our understanding of the underlying physics before the LISA mission commences, the expected BHBH detection rate is unlikely to change significantly. In contrast, the NSNS and BHNS rates are very sensitive to changes in binary physics assumptions. Therefore, once LISA flies and we know the actual number of detections, we can compare to each model and possibly provide some constraint on binary evolution physics.

Here is some attempted reasoning about the trends:

\textbf{BHBH:} The major standout for the BHBHs is the optimistic. In this case Hertzsprung gap donors are allowed to survive CE and this is because many BHs are formed in this way rather than any effect on detectability. Can also comment on lower(higher) max NS mass being good(bad). Kicks same as BHNSs but less pronounced.

\textbf{BHNS:} Whole bunch of trends here. The decrease with increasing $\beta$ seems to be due to fewer BHNSs surviving CE events (since envelope is harder to eject). Floor explains this in her paper and also references Kruckow. optimistic higher same as BHBH. Rapid higher because BHBHs become BHNSs instead. Maximum neutron star masses affect things because: lower means more things become BH instead of NS hence fewer BHNS and vice versa. PISN doesn't affect this mass range. Lower kicks means fewer disruptions. No BH kick lower again because NS kick is main factor at that point. \todo{this leaves E, F, G todo}

\textbf{NSNS:} \todo{Why are B, C, D opposite?}. E is low because that's the main NSNS formation channel. CE follows nice trend (see Floor's about CE channels). Optimistic slightly up because duh. \todo{weirdly basically no change in rapid and max NS masses}. Lower kicks great as usual. No BH kicks no effect duh.

@@

\begin{table*}[htb]
    \centering
    \caption{The number of detectable binaries in a 4 and 10 year LISA mission for the 15 different model variations and each DCO type. Each value shows the median and the 90\% confidence interval.}
    \begin{tabular}{c|lll|lll}
        \hline
        \multirow{2}{*}{Model} & \multicolumn{3}{c|}{LISA (4 year)} & \multicolumn{3}{c}{LISA (10 year)} \\ \cline{2-7}
         & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} \\
        \hline
        A & \confinv{25.9}{11.1}{13.6} & \confinv{26.7}{11.9}{14.8} & \confinv{11.3}{6.4}{8.0} & \confinv{42.0}{17.3}{17.3} & \confinv{44.5}{17.8}{20.7} & \confinv{19.3}{8.0}{9.7}\\
        B & \confinv{27.7}{12.6}{12.6} & \confinv{12.0}{6.0}{7.5} & \confinv{4.7}{2.7}{3.3} & \confinv{42.7}{17.6}{17.6} & \confinv{19.5}{7.5}{10.5} & \confinv{8.0}{3.3}{4.0}\\
        C & \confinv{20.6}{9.4}{11.3} & \confinv{6.2}{2.8}{3.4} & \confinv{5.5}{2.7}{4.1} & \confinv{33.8}{13.1}{13.1} & \confinv{10.1}{3.9}{3.9} & \confinv{9.6}{4.1}{4.1}\\
        D & \confinv{20.0}{8.3}{10.0} & \confinv{3.2}{1.6}{1.6} & \confinv{14.6}{8.3}{10.4} & \confinv{31.7}{11.7}{11.7} & \confinv{5.1}{1.9}{2.2} & \confinv{22.9}{10.4}{12.5}\\
        E & \confinv{24.4}{12.2}{14.7} & \confinv{6.8}{3.8}{3.8} & \confinv{0.2}{0.1}{0.1} & \confinv{41.5}{17.1}{17.1} & \confinv{10.6}{3.8}{5.3} & \confinv{0.4}{0.1}{0.2}\\
        F & \confinv{20.0}{10.0}{10.0} & \confinv{11.5}{6.6}{8.2} & \confinv{8.1}{3.6}{5.4} & \confinv{30.0}{12.0}{14.0} & \confinv{19.7}{9.9}{9.9} & \confinv{13.6}{5.4}{6.3}\\
        G & \confinv{25.7}{9.9}{13.8} & \confinv{23.3}{10.6}{10.6} & \confinv{26.5}{15.1}{18.9} & \confinv{39.5}{13.8}{15.8} & \confinv{36.1}{12.7}{17.0} & \confinv{45.4}{18.9}{22.7}\\
        H & \confinv{63.6}{31.8}{47.7} & \confinv{37.4}{20.8}{20.8} & \confinv{14.7}{8.4}{10.5} & \confinv{103.4}{39.8}{55.7} & \confinv{62.3}{24.9}{29.1} & \confinv{25.2}{12.6}{12.6}\\
        I & \confinv{21.7}{10.9}{13.0} & \confinv{56.8}{31.6}{31.6} & \confinv{11.0}{5.5}{6.9} & \confinv{32.6}{13.0}{15.2} & \confinv{94.7}{37.9}{44.2} & \confinv{17.9}{8.3}{8.3}\\
        J & \confinv{32.6}{16.3}{19.6} & \confinv{20.3}{9.0}{13.5} & \confinv{11.0}{6.3}{7.8} & \confinv{52.2}{19.6}{22.8} & \confinv{33.8}{11.3}{15.8} & \confinv{18.8}{7.8}{9.4}\\
        K & \confinv{21.4}{9.7}{11.7} & \confinv{30.6}{13.6}{17.0} & \confinv{11.5}{6.6}{8.2} & \confinv{35.0}{13.6}{15.6} & \confinv{51.1}{20.4}{23.8} & \confinv{19.8}{8.2}{9.9}\\
        L & \confinv{24.7}{9.9}{14.8} & \confinv{26.7}{14.8}{17.8} & \confinv{11.3}{6.4}{8.0} & \confinv{42.0}{14.8}{17.3} & \confinv{44.5}{17.8}{20.7} & \confinv{19.3}{8.0}{9.7}\\
        M & \confinv{29.7}{13.2}{16.5} & \confinv{70.6}{39.2}{39.2} & \confinv{20.4}{11.7}{11.7} & \confinv{46.1}{19.8}{23.1} & \confinv{109.8}{39.2}{54.9} & \confinv{32.1}{14.6}{17.5}\\
        N & \confinv{31.6}{15.8}{19.7} & \confinv{117.7}{58.8}{73.6} & \confinv{51.8}{29.6}{29.6} & \confinv{51.3}{23.7}{23.7} & \confinv{191.3}{73.6}{103.0} & \confinv{81.4}{37.0}{44.4}\\
        O & \confinv{30.2}{15.1}{18.9} & \confinv{96.7}{48.3}{60.4} & \confinv{11.4}{6.5}{8.1} & \confinv{49.1}{22.7}{22.7} & \confinv{157.1}{60.4}{84.6} & \confinv{19.5}{9.7}{9.7}\\
        \hline
    \end{tabular}
    \label{tab:detection_rates}
\end{table*}

@@

We illustrate the distribution of detectable DCOs on the sensitivity curve in Figure~\ref{fig:dcos_on_sc}. The shape of the distributions are bounded by the edge of the Milky Way and the scarcity of nearby DCOs. We indicate this with lines of constant distance that form as a boundary for the majority of DCOs, where the remaining scatter is due to variations in mass since lines are plotted for the average chirp mass of each respective DCO. The top edge is not a hard cutoff but is present because for any given distance there is a lower probability of drawing higher frequencies (since binaries spend more time at lower frequencies).

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{dcos_on_sc.png}
    \caption{Density distribution of detectable DCOs plotted over the LISA sensitivity curve \citep{Robson+2019}. We plot the total signal of binaries at their dominant frequency $n f_{\rm orb}$, such that $n$ is the harmonic that produces the most relative gravitational wave luminosity ($n = 2$ for circular binaries). If the density of points is below our lowest contour (2\%) then we plot the points as scatter points, where their sizes corresponds to their STROOPWAFEL weights. In order to explain the edges of the distribution, we additionally plot lines of constant distance. The dotted lines show the signal for a circular binary of average chirp mass, whilst the dashed line shows the signal for a binary with $e = 0.9$}
    \label{fig:dcos_on_sc}
\end{figure*}

\subsection{Distributions for the fiducial model}\label{sec:fiducial_distributions}

@@
